---
title: "Optimization with boundary layer conductance"
author: "Remko Duursma"
date: "Monday, June 16, 2014"
output:
  html_document:
    fig_caption: yes
---

```{r, echo=FALSE}
library(knitr)
opts_chunk$set(echo=FALSE, warning=FALSE, error=FALSE, message=FALSE)
```



### Exploring the Penman-Monteith equation

Here, a couple of key figures using the Penman-Monteith equation. Leaf temperature is calculated by closing the energy balance (numerically), taking into all temperature effects. For the following figures, wind speed (and hence boundary layer conductance) and stomatal conductance were inputs, and transpiration and leaf temperature are simulated.


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(plantecophys)
library(magicaxis)
library(pander)

windlabel <- expression(Wind~speed~~(m~s^-1))


gslow <- 0.05
gshigh <- 0.5

Winds <- exp(seq(log(0.1), log(10), length=25))
```

These simulations are with `plantecophys` package version `r packageVersion("plantecophys")`.


```{r echo=FALSE, message=FALSE, fig.cap="Figure 1. Leaf temperature calculated from energy balance, given a constant low or high stomatal conductance. Note how wind speed increases leaf temperature when gs is high, but decreases it when it is low. Monteith (1965) showed a similar figure."}

# Calculate Tleaf given known gs, from energy balance
rlow <- sapply(Winds, function(x)plantecophys:::FindTleaf(gs=gslow, Wind=x, Tair=25))
rhigh <- sapply(Winds, function(x)plantecophys:::FindTleaf(gs=gshigh, Wind=x, Tair=25))

plot(log10(Winds), rlow, ylim=c(20,30), type='l', col="red", axes=FALSE,
     xlab=windlabel, ylab=expression(T[leaf]~~(degree*C)))
points(log10(Winds), rhigh, type='l', col="blue")

abline(h=25)
magaxis(side=1, unlog=1)
axis(2)
box()
legend("topright", legend=sapply(c(bquote(g[s] == .(gslow)),
                          bquote(g[s] == .(gshigh)),
                          bquote(T[air])
                          ),as.expression), lty=1, col=c("red","blue","black"),
       cex=0.9)
```



```{r echo=FALSE, fig.cap="Figure 2. Leaf transpiration estimated from stomatal conductance and wind speed, using Penmon-Monteith and solved leaf temperature. At low wind speed, the low boundary layer conductance leads to a curved relationship between E and gs. Note that at low gs, E is higher in low wind speed. This is due to increased leaf temperature as a result of low conductance to heat, which increases the leaf-to-air VPD. At high wind speed, we see the approximate E = gs*VPD relationship."}

# Calculate ELEAF given known gs; calculate Tleaf from energy balance.
f <- function(w,gs,...){
  
  tleaf <- plantecophys:::FindTleaf(gs=gs, Wind=w, Tair=25,VPD=1)
  flux <- plantecophys:::LeafEnergyBalance(Tleaf=tleaf, Wind=w, Tair=25, gs=gs, 
                                           VPD=1,
                                           returnwhat="fluxes")
  flux$Tleaf <- tleaf

return(flux)
}

# Wind speed and E/gs
gss <- seq(0.02, 0.5, length=25)
windlow <- 0.1
windhigh <- 10

wlow <- do.call(rbind, lapply(gss, function(x)f(windlow,x)))
whigh <- do.call(rbind, lapply(gss, function(x)f(windhigh,x)))

plot(gss, wlow$ELEAFeb, type='l', col="red", ylim=c(0,5),xlim=c(0,0.6),
     xlab=expression(g[s]~~(mol~m^-2~s^-1)),
     ylab=expression(E[leaf]~~(mmol~m^-2~s^-1)))

points(gss, whigh$ELEAFeb, type='l', col="blue")

legend("topleft", legend=sapply(c(bquote(wind == .(windlow)),
                                   bquote(wind == .(windhigh))),as.expression), 
       lty=1, col=c("red","blue"), cex=0.9)
```




```{r echo=FALSE, fig.cap="Figure 3. Test of the approximate relationship E = gs*VPD (note molar units), with varying wind speed and at low and high stomatal conductance. For this simulation, VPD was set to 0.01 mol mol-1 (ca. 1 kPa). As we know, at high wind speeds (when leaves are 'well coupled'), the ratio E/gs should then converge to VPD, as it does. Consistent with the above figures, there is an interaction between wind speed and gs."}
rlow <- do.call(rbind,lapply(Winds, function(x)f(x, gs=gslow)))
rhigh <- do.call(rbind,lapply(Winds, function(x)f(x, gs=gshigh)))

plot(log10(Winds), rlow$ELEAFeb/gslow/1000,  type='l', col="red", ylim=c(0,0.02),
     axes=FALSE,  xlab=windlabel, ylab=expression(E[leaf]/g[s]~~(mol~mol^-1)))
points(log10(Winds), rhigh$ELEAFeb/gshigh/1000, type='l', col="blue")
abline(h=0.01)

magaxis(side=1, unlog=1)
axis(2)
box()
legend("topright", legend=sapply(c(bquote(g[s] == .(gslow)),
                                   bquote(g[s] == .(gshigh)),
                                   bquote(VPD)), as.expression),
       lty=1, col=c("red","blue","black"))

f <- lapply(Winds, function(x)FARAO2(Wind=x, energybalance=TRUE))
f <- do.call(rbind,f)
```




### FARAO with energy balance

Here, I show some simulations with FARAO (FARquhar And Optimization). A new implementation follows Buckley et al. (2014). Using the leaf gas exchange model (that takes Ci, Tair - and other drivers of course - and calculates A, E, gs and Tleaf), lambda was calculated numerically with,

$$\frac{dA}{dE} == \frac{dA / dC_i}{dE / dC_i}$$

This was done by calculating A and E at a given Ci, then adding a very small number, and calculating $dA = A(C_i+d) - A(C_i)$. Then, the Ci at which the calculated lambda was equal to a preset value was found by optimization.


```{r eval=FALSE, echo=FALSE, fig.cap="Figure 4. VPD response of GS at constant Tair, using the full numerical optimization with energy balance. At lower wind speed, a higher gs is achieved at a given leaf-to-air VPD, but this VPD is also much higher because Tleaf is increased under low gbl. Note that at high VPD, the numerical solution does not converge (this means there is no optimal gs in that region, cf. Buckley et al 2014).",cache=F}
vpds <- seq(1, 3.5, length=25)
windlow <- 0.4
windhigh <- 10
runcon <- FARAO2(energybalance=FALSE, VPD=vpds, Tleaf=25)
run1 <- FARAO2(energybalance=TRUE, VPD=vpds, Wind=windlow, Tair=25)
run2 <- FARAO2(energybalance=TRUE, VPD=vpds, Wind=windhigh, Tair=25)

run1 <- run1[is.finite(run1$ELEAF),]
run2 <- run2[is.finite(run2$ELEAF),]

f2 <- function(x)10^-3*x$ELEAF/(x$VPD/101)
runcon$GSinf <- f2(runcon)
run1$GSinf <- f2(run1)
run2$GSinf <- f2(run2)

f <- function(x, vpdname="VPD")x$ALEAF/(x$Ca*sqrt(x[,vpdname]))
runcon$bb <- f(runcon)
run1$bb <- f(run1)
run2$bb <- f(run2)
run1$bb2 <- f(run1,"VPDleaf")
run2$bb2 <- f(run2, "VPDleaf")

with(runcon, plot(VPD, GS, type='l', ylim=c(0,0.16),
                  xlim=c(1,5),
                  xlab="VPD (air) (kPa)",
                  ylab=expression(g[s]~~(mol~m^-2~s^-1))))
with(run1, points(VPDleaf, GS, type='l', col="red"))
with(run2, points(VPDleaf, GS, type='l', col="blue"))
legend("bottomleft", c("No energy balance","Wind = 0.4", "Wind = 10"),
       lty=1, col=c("black","red","blue"))

with(runcon, plot(VPD, ALEAF/ELEAF, type='l', ylim=c(0,10),
                  xlim=c(1,5),
                  xlab="VPD (air) (kPa)",
                  ylab=expression(g[s]~~(mol~m^-2~s^-1))))
with(run1, points(VPDleaf, ALEAF/ELEAF, type='l', col="red"))
with(run2, points(VPDleaf, ALEAF/ELEAF, type='l', col="blue"))
legend("bottomleft", c("No energy balance","Wind = 0.4", "Wind = 10"),
       lty=1, col=c("black","red","blue"))



with(runcon, plot(bb, GS, xlim=c(0,0.04), pch=19))
with(run1, points(bb, GS, col="red", pch=19))
with(run2, points(bb, GS, col="blue", pch=19))

with(runcon, plot(bb, GS, xlim=c(0,0.04), pch=19))
with(run1, points(bb, GSinf, col="red", pch=19))
with(run2, points(bb, GSinf, col="blue", pch=19))


```


```{r, warning=FALSE, error=FALSE, echo=FALSE, fig.cap="Figure 5. PPFD, Tair, VPD and Ca were varied randomly, gs was calculated with FARAO, either without energy balance (assuming Tleaf=Tair), or solving for Tleaf depending on wind speed. Here solved gs is plotted against the simple index, using the leaf VPD (as we usually assume is most appropriate). Clearly, g1 has increased (see Table further below).", cache=F}

n <- 101
set.seed(1111)
pars <- runif(n, 200,1000)
vpds <- runif(n,1,3.5)
tairs <- runif(n,15,28)
cas <- runif(n, 380, 500)
windlow <- 0.4
windhigh <- 10


rcon <- FARAO2(energybalance=FALSE, VPD=vpds, PPFD=pars, Tleaf=tairs, Ca=cas)
r1 <- FARAO2(energybalance=TRUE, VPD=vpds, Tair=tairs, PPFD=pars, Ca=cas, Wind=windlow)
r2 <- FARAO2(energybalance=TRUE, VPD=vpds, Tair=tairs, PPFD=pars, Ca=cas, Wind=windhigh)
r1 <- r1[is.finite(r1$ELEAF),]
r2 <- r2[is.finite(r2$ELEAF),]

f <- function(x, vpdname="VPD")x$ALEAF/(x$Ca*sqrt(x[,vpdname]))
rcon$bb <- f(rcon)
r1$bb <- f(r1)
r2$bb <- f(r2)
r1$bb2 <- f(r1,"VPDleaf")
r2$bb2 <- f(r2, "VPDleaf")

# inferred GS using air VPD (this is what we might do for eddy data or when gbl is unknown)
f2 <- function(x)10^-3*x$ELEAF/(x$VPD/101)
rcon$GSinf <- f2(rcon)
r1$GSinf <- f2(r1)
r2$GSinf <- f2(r2)


with(rcon, plot(bb, GS, pch=19, ylim=c(0,0.25), xlim=c(0,0.04),
                  xlab=expression(A/C[a]*sqrt(D[leaf])), 
                  ylab=expression(Optimized~g[s]~~(mol~m^-2~s^-1))))
with(r1, points(bb2, GS, pch=19, col="red"))
with(r2, points(bb2, GS, pch=19, col="blue"))
legend("topleft", c("No energy balance","Wind = 0.4", "Wind = 10"),
       lty=1, col=c("black","red","blue"))

abline(lm(GS ~ bb, data=rcon))
abline(lm(GS ~ bb2, data=r1), col="red")
abline(lm(GS ~ bb2, data=r2), col="blue")

```

```{r eval=FALSE}
with(rcon, plot(VPD, ALEAF/ ELEAF, xlim= c(0,5), ylim=c(0,15)))
with(r1, points(VPD, ALEAF/ELEAF, col="red"))
with(r2, points(VPD, ALEAF/ELEAF, col="blue"))

with(rcon, plot(VPD, ALEAF/ ELEAF, xlim= c(0,5), ylim=c(0,15)))
with(r1, points(VPDleaf, ALEAF/ELEAF, col="red"))
with(r2, points(VPDleaf, ALEAF/ELEAF, col="blue"))
```



```{r, echo=FALSE, eval=FALSE, fig.cap="Figure 6. If we instead calculate the simple index using the air VPD, the differences are smaller. Still, at low wind speed, g0 and/or g1 is higher. This makes sense because with low gbl, E increases less quickly with gs, so that if lambda is constant, the stomata can open a bit more at the same cost of water."}

with(rcon, plot(bb, GS, pch=19, ylim=c(0,0.25), xlim=c(0,0.04),
                  xlab=expression(A/C[a]*sqrt(D[air])), 
                  ylab=expression(g[s]~~(mol~m^-2~s^-1))))
with(r1, points(bb, GS, pch=19, col="red"))
with(r2, points(bb, GS, pch=19, col="blue"))
legend("topleft", c("No energy balance","Wind = 0.4", "Wind = 10"),
       lty=1, col=c("black","red","blue"))

abline(lm(GS ~ bb, data=rcon))
abline(lm(GS ~ bb, data=r1), col="red")
abline(lm(GS ~ bb, data=r2), col="blue")
```


```{r, echo=FALSE, fig.cap="Figure 7. This figure shows gs inferred from leaf transpiration, assuming infinite boundary layer conductance (gs = E/Dair), and Dair is used in the stomatal index. This is what we would do for field conditions when we don't know the boundary layer conductance and are happy to assume that 'the leaves are well-coupled'. It shows that the three relationships now almost collapse: the errors associated with assuming infinite boundary layer conductance cancel out."}

with(rcon, plot(bb, GSinf, pch=19, xlim=c(0,0.04),  ylim=c(0,0.2), 
                  xlab=expression(A/C[a]*sqrt(D[air])), 
                  ylab=expression(Inferred~g[s]~~(mol~m^-2~s^-1))))
with(r1, points(bb, GSinf, pch=19, col="red"))
with(r2, points(bb, GSinf, pch=19, col="blue"))
abline(lm(GSinf ~ bb, data=rcon))
abline(lm(GSinf ~ bb, data=r1), col="red")
abline(lm(GSinf ~ bb, data=r2), col="blue")

legend("topleft", c("No energy balance","Wind = 0.4", "Wind = 10"),
       lty=1, col=c("black","red","blue"))

```



I also fit g1 with the Medlyn et al. 2011 model using non-linear regression (and assuming g0=0). The table below shows two estimates: the first is using leaf-to-air VPD, and fitting *actual* optimized gs. The second estimate is using the air VPD, and using the *inferred* gs from Eleaf (by gs = E/VPD).


```{r, echo=FALSE, results='asis'}


fitg1 <- function(x, vpdname, gsname="GS", g0=TRUE ){
  x$VPD <- x[,vpdname]
  x$gs <- x[,gsname]
  if(g0){
    fit <- nls(gs ~ g0 + 1.6*(1+g1/sqrt(VPD))*ALEAF/Ca, data=x,
             start=list(g0=0, g1=4))
    return(coef(fit))
  } else {
    fit <- nls(gs ~ 1.6*(1+g1/sqrt(VPD))*ALEAF/Ca, data=x,
             start=list(g1=4))
    return(coef(fit))
  }
  
}


df <- data.frame(g1_estimate1=c(fitg1(rcon, "VPD", g0=F),
                      fitg1(r2, "VPDleaf", g0=F),
                      fitg1(r1, "VPDleaf", g0=F)),
                 g1_estimate2=c(fitg1(rcon, "VPD", g0=F),
                      fitg1(r2, "VPD", "GSinf", g0=F),
                      fitg1(r1, "VPD", "GSinf", g0=F)))
rownames(df) <- c("Infinite gbl","Wind = 10", "Wind = 0.4")

pander(df)

```







### Caveats

* I have not yet implemented the effect of boundary layer conductance on photosynthesis, but this is probably only a small correction.
* The full coupled numerical optimization does not always find a solution, this is probably not because the code is bad, but because in quite a lot of cases there may not be an optimal solution (see Buckley et al. 2014). It especially does not converge when VPD is very high or low, or very high at lowish temperature. In the above, about 30-40% of the cases I could not find a solution, I hope this does not skew the results too much!
* Another caveat is that I don't really understand why this comes out the way it does!




```{r echo=FALSE, eval=FALSE, warning=FALSE}
# This is to check it works; here we found a discnontinuity before.
windlow <- 0.5
windhigh <- 10

Cis <- seq(100,375, length=101)

rlow <- PhotosynEB(Ci=Cis, Wind=windlow)
rhigh <- PhotosynEB(Ci=Cis, Wind=windhigh)
rcon <- Photosyn(Ci=Cis)

with(rcon, plot(Ci, ALEAF, type='l'))
with(rlow, points(Ci, ALEAF, type='l', col="red"))
with(rhigh, points(Ci, ALEAF, type='l', col="blue"))

with(rcon, plot(Ci, ELEAF, type='l'))
with(rlow, points(Ci, ELEAF, type='l', col="red"))
with(rhigh, points(Ci, ELEAF, type='l', col="blue"))

with(rcon, plot(Ci, Tleaf, type='l'))
with(rlow, points(Ci, Tleaf, type='l', col="red"))
with(rhigh, points(Ci, Tleaf, type='l', col="blue"))
```






```{r, eval=FALSE}


w <- seq(0.2, 10, length=25)

r <- FARAO2(energybalance=TRUE, Wind=w, Ca=400, Tair=25, VPD=2, Wleaf=0.05)

f2 <- function(x)10^-3*x$ELEAF/(x$VPD/101)

g1 <- function(x, vpdname="VPD", gsname="GS"){
  with(x, sqrt(x[,vpdname])*((x[,gsname]/1.6)*(Ca/ALEAF) -1))
}

r$g1_VPD <- g1(r, "VPD")
r$g1_VPDleaf <- g1(r, "VPDleaf")
r$GSinf <- f2(r)
r$g1_VPD_gsinf <- g1(r, "VPD", "GSinf")

with(r, plot(Gbh, g1_VPD, type='l', xlim=c(0,6), ylim=c(0,5)))
with(r, points(Gbh, g1_VPDleaf, type='l', lty=5))

with(r, points(Gbh, g1_VPD_gsinf, type='l', lwd=2))





getg1 <- function(w, dfr){
  
  rcontrol <- FARAO2(energybalance=FALSE, Ca=dfr$Ca, Tleaf=dfr$Tair, 
              VPD=dfr$VPD, PPFD=dfr$PPFD)
  r <- FARAO2(energybalance=TRUE, Wind=w, Ca=dfr$Ca, Tair=dfr$Tair, 
              VPD=dfr$VPD, Wleaf=0.05, PPFD=dfr$PPFD)
  
  fitg1 <- function(x, vpdname, gsname="GS"){
    x$VPD <- x[,vpdname]
    x$gs <- x[,gsname]
    fit <- nls(gs ~ 1.6*(1+g1/sqrt(VPD))*ALEAF/Ca, data=x,
             start=list(g1=4))
    return(coef(fit))
  }
  r$GSinf <- with(r, 10^-3*ELEAF/(VPD/101))

  l <- list(data=r, datacontrol=rcontrol, g1=data.frame(g1_0=fitg1(rcontrol,"VPD"),
                              g1_VPDair=fitg1(r, "VPD"),
                              g1_VPDleaf=fitg1(r, "VPDleaf"),
                              g1_VPDair_GSinf=fitg1(r, "VPD", "GSinf")
                              ))
  class(l) <- "gblrun"
  return(l)
  
}

plot.gblrun <- function(x,...){

  
  with(x$data, plot(ALEAF/(sqrt(VPD)*Ca), GS, pch=19,
                    ylim=c(0, max(GS)), xlim=c(0,0.05)))
  with(x$datacontrol, points(ALEAF/(sqrt(VPD)*Ca), GS, pch=19, col="darkgrey"))
  
  with(x$data, plot(ALEAF/(sqrt(VPDleaf)*Ca), GS, pch=19,col="blue",
                    ylim=c(0, max(GS)), xlim=c(0,0.05)))
  with(x$datacontrol, points(ALEAF/(sqrt(VPD)*Ca), GS, pch=19, col="darkgrey"))

  with(x$data, plot(ALEAF/(sqrt(VPD)*Ca), GSinf, pch=19,col="forestgreen",
                    ylim=c(0, max(GS)), xlim=c(0,0.05)))
  with(x$datacontrol, points(ALEAF/(sqrt(VPD)*Ca), GS, pch=19, col="darkgrey"))
  
}
  
vpds <- c(seq(1, 3.5, by=0.5))
rhs <- seq(20,60,by=5)
cas <- c(380,600)
tairs <- 22
ppfds <- c(400,800,1200)
dfr <- expand.grid(RH=rhs, Tair=tairs, Ca=cas, PPFD=ppfds)
dfr$VPD <- RHtoVPD(dfr$RH,dfr$Tair)

g <- getg1(0.8, dfr)


w <- seq(0.5, 5, length=10)
res <- lapply(w, function(x)getg1(x,dfr))

g1df <- do.call(rbind, lapply(res, "[[", "g1"))
g1df$Wind <- w

with(g1df, plot(Wind, g1_0, type='l', lwd=2, ylim=c(1.5,4.2)))
with(g1df, points(Wind, g1_VPDair, type='l', col="red"))
with(g1df, points(Wind, g1_VPDleaf, type='l', col="blue"))
with(g1df, points(Wind, g1_VPDair_GSinf, type='l', col="forestgreen"))

d <- do.call(rbind, lapply(res, "[[", "data"))
d$bb_VPD <- with(d, ALEAF / (sqrt(VPD)*Ca))
d$bb_VPDleaf <- with(d, ALEAF / (sqrt(VPDleaf)*Ca))

palette(rainbow(10))
with(d, plot(bb_VPD, GS, pch=15, col=as.factor(Wind), ylim=c(0,0.2), xlim=c(0,0.05)))
with(d, plot(bb_VPDleaf, GS, pch=15, col=as.factor(Wind), ylim=c(0,0.2), xlim=c(0,0.05)))
with(d, plot(bb_VPD, GSinf, pch=15, col=as.factor(Wind), ylim=c(0,0.2), xlim=c(0,0.05)))


```





